Evaluating an attempt to restore summer fire in the Northern Great Plains

### 2025-04-21 SpringSummerSeverity/paper/SupplementalInformation.Rmd
Spring/summer seasonality R1 update

@@

# Remote sensing

## Sentinel-2

@@

\clearpage
## Landsat

@@

# Analysis

@@

## Remotely sensed imagery

@@

## Weather data

@@

seasons <- tibble(season = c('spring', 'summer'), 
                 start = c('04-20', '07-20'), 
                 end = c('06-10', '09-10') ) %>%
          mutate(start = as.Date(start, '%m-%d'), 
                 end = as.Date(end, '%m-%d'))

@@

# Hourly data 
  WxData$HourlyWx <- 
    read_xlsx('./data/CGREC_weather.xlsx', 'hourly') %>%
    filter(between(Hour, 1000, 1700)) %>%
    unite( c(Year, Month, Day), col = 'date', sep = '-', remove = F) %>%
    unite( c(Month, Day), col = 'day', sep = '-', remove = F) %>%
    mutate(day = as.Date(day, '%m-%d'), 
           season = case_when(
             between(day, seasons$start[1], seasons$end[1]) ~'Spring', 
             between(day, seasons$start[2], seasons$end[2]) ~'Summer',
             TRUE ~ NA      
           ) ) %>%
    mutate(e = 6.11 * (10 ^ ( (7.5 * DewPoint)/ (237.3 + DewPoint) ) ), 
           es = 6.11 * (10 ^ ( (7.5 * AirTemp)/ (237.3 + AirTemp) ) ), 
           VPD = es - e) 
# Get burn day weather
  WxData$BurnDayWx <-
    HourlyWx  %>% 
    filter(date %in% BurnDays$date) %>%
    group_by(Year, date, season) %>%
    summarise_at(.vars = vars(c(DewPoint, RelHum, WindSpeed,VPD)),
                 .funs = c("mean")) %>%
    ungroup() %>%
    pivot_longer(names_to = 'variable', 
                 values_to = 'value', 
                 cols = DewPoint:VPD) %>%
    mutate(variable = recode(variable, 
                             'DewPoint' = 'Dew Point (°C)', 
                             'RelHum' = 'Relative Humidity (%)', 
                             'VPD' = 'Vapor Pressure Deficit', 
                             'WindSpeed' = 'Wind speed (m/s)') , 
           Year = as.character(Year)) 
\end{lstlisting}

>>>>

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE)
pacman::p_load(tidyverse, xtable, wesanderson)
  BuGnPal = c( "#67A9CF" ,"#3690C0" ,"#02818A", "#016450")
  FFgrad = wes_palette("FantasticFox1")[c(3,2,4,5)]
  load('../data/NoFireIndices.Rdata')
  # load('../data/BurnIndices.Rdata')
  # load('../data/NoFireTrends.Rdata')
  load('../data/WxData.Rdata')
# some necessary things
  sys_yr <- 
    Sys.Date() %>%
      as.POSIXct(., '%Y-%M-%d') %>%
      year() 
  seasons <- 
    tibble(season = c('Spring', 'Summer'), 
           start = c(as.Date(paste0(sys_yr, '-04-25')), as.Date(paste0(sys_yr, '-08-07'))), 
           end = c(as.Date(paste0(sys_yr, '-05-25')), as.Date(paste0(sys_yr, '-08-27'))))
```

# Additional data and results

## Seasonal trends

```{r ndvi_trend, eval = TRUE, fig.cap="Seasonal trends in Landsat-derived NDVI for the study period. Green shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 4, fig.width = 8}
NoFireIndices %>%
  group_by(ImageDate, pasture) %>%
    summarise(Mean = mean(ndvi), 
              SEM = sd(ndvi)/sqrt(n()), 
              .groups = 'drop') %>%
  mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
         Year = lubridate::year(ImageDate), 
         Year = as.factor(Year), 
         SampleDate = format(ImageDate, '%m-%d'), 
         SampleDate = as.Date(SampleDate, '%m-%d') ) %>%
  filter(Year != 2016) %>%
ggplot(aes(x = SampleDate, 
           color = Year, 
           fill = Year, 
           group = interaction(pasture, Year))) + theme_bw(14) +
  geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf), 
              fill='#A1D99B', alpha=0.5) +
  geom_smooth(aes(y = Mean, group = Year),
              se = T,
              level = 0.99) +
  geom_errorbar(aes(ymin = Mean - SEM, 
                    ymax = Mean + SEM),
                alpha = 0.5) +
  geom_line(aes(y = Mean),
             alpha = 0.5) +
  labs(x = 'Image date', y = 'NDVI') + 
    scale_fill_manual(values = FFgrad) + 
  scale_color_manual(values = FFgrad) + 
  scale_x_date( date_breaks = 'month', 
                date_labels = '%b-%d', 
                minor_breaks = NULL) 
```

```{r precip_trend, eval = TRUE, fig.cap="Historical precipitation data for the entire 42-year period. Thick black line shows smoothed 42-yr trend, gray lines individual years, and colored lines study years. Green shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 4, fig.width = 8}
  WxData$Rainfall %>% 
  mutate(day = day + years(1)) %>%
    filter(month(date) %in% c(04:09)) %>%
    separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
    group_by(Year) %>%
    mutate(CumPrecip = cumsum(Rainfall)) %>%
    ggplot(aes(x = day)) + theme_bw(14) +
    geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf),
              fill='#A1D99B', alpha=0.5) +
    geom_line(aes(y = CumPrecip, group = Year),
              color = 'grey40', 
              alpha = 0.5) +
    geom_smooth(aes(y = CumPrecip),
                se = F,
                color = 'black',
                size = 1,
                level = 0.99) +
    geom_line(data = . %>%
                filter(as.numeric(Year) %in% c(2017:2020)) , 
              aes(y = CumPrecip, 
                  color = Year,
                  group = Year),
              size = 2) +
    labs(x = 'Day of year', y = 'Cumulative rainfall (mm)') + 
    scale_fill_manual(values = FFgrad) + 
    scale_color_manual(values = FFgrad) + 
    scale_x_date( date_breaks = 'month', 
                  date_labels = '%b-%d', 
                  minor_breaks = NULL)
```

\clearpage

## $z$ scores for historical data by burn season

```{r historical_z, eval = TRUE, results='asis'}
load('../data/z_tab.Rdata')
 z_tab$hist %>%
  xtable(caption = 'Seasonal deviations in fuel conditions and fire weather for spring and summer burn seasons, relative to two decades (1990-1999, 2012-2022) as presented in Fig. 4. Anomalies determined by $z$ scores (parentheses) and categorized as above, below, or within 1 standard deviation of the mean for the given historical range. 2018 and 2019 were the years in which no summer burns were conducted; all spring burns were conducted in all years') %>%
   print(include.rownames = F, comment = F)
```

# Script

@@

## Remote sensing

### Sentinel-2

@@

\clearpage
### Landsat

@@

## Analysis

@@

### Remotely sensed imagery

@@

### Weather data

@@

  sys_yr <- 
    Sys.Date() %>%
      as.POSIXct(., '%Y-%M-%d') %>%
      year() 
  seasons <- 
    tibble(season = c('Spring', 'Summer'), 
           start = c(as.Date(paste0(sys_yr, '-04-20')), 
                     as.Date(paste0(sys_yr, '-07-20'))), 
           end = c(as.Date(paste0(sys_yr, '06-10')), 
                   as.Date(paste0(sys_yr, '09-10'))))

@@

# Hourly data 
  WxData$HourlyWx <- 
    read_xlsx('./data/CGREC_weather.xlsx', 'hourly') %>%
    filter(between(Hour, 1000, 1700)) %>%
    unite( c(Year, Month, Day), col = 'date', sep = '-', remove = F) %>%
    unite( c(Month, Day), col = 'day', sep = '-', remove = F) %>%
    mutate(day = as.Date(day, '%m-%d'), 
           season = case_when(
             between(day, seasons$start[1], seasons$end[1]) ~'Spring', 
             between(day, seasons$start[2], seasons$end[2]) ~'Summer',
             TRUE ~ NA  ) ) %>%
    mutate(e = 6.11 * (10 ^ ( (7.5 * DewPoint)/ (237.3 + DewPoint) ) ), 
           es = 6.11 * (10 ^ ( (7.5 * AirTemp)/ (237.3 + AirTemp) ) ), 
           VPD = es - e) 
# Get burn day weather
  WxData$BurnDayWx <-
    HourlyWx  %>% 
    filter(date %in% BurnDays$date) %>%
    group_by(Year, date, season) %>%
    summarise_at(.vars = vars(c(DewPoint, RelHum, WindSpeed,VPD)),
                 .funs = c("mean")) %>%
    ungroup() %>%
    pivot_longer(names_to = 'variable', 
                 values_to = 'value', 
                 cols = DewPoint:VPD) 
\end{lstlisting}

### Statistical analysis

\textsf{R} script for statistical analysis is organized by the manuscript figure for which the tests support.

**Figure 1**

# A: Test NDVI in study years across seasons 
  # Wrangle NDVI data from adjacent unburned pastures
    ndvi_dat <-
      NoFireIndices %>%
        group_by(ImageDate, pasture) %>%
        summarise(NDVI = median(ndvi) , 
                  .groups = 'drop') %>%
        mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
               Year = lubridate::year(ImageDate), 
               Year = as.factor(Year), 
               SampleDate = format(ImageDate, '%m-%d'), 
               SampleDate = as.Date(SampleDate, '%m-%d'), 
               season = case_when(
                 between(SampleDate, seasons$start[1], seasons$end[1]) ~ 'Spring',
                 between(SampleDate, seasons$start[2], seasons$end[2]) ~ 'Summer', 
                 TRUE ~ NA ) ) %>%
        filter(!is.na(season)) %>%
      select(pasture, Year, season, NDVI)
  # Fit mixed-effect models
    m0 <- lme4::lmer(NDVI ~ 1 + (1|Year), REML=FALSE, data = ndvi_dat) 
    m1 <- lme4::lmer(NDVI ~ season + (1|Year), REML=FALSE, data = ndvi_dat)
  # Compare models for significance of season term
    anova(m0, m1)

# B: Test for difference between seasons
  # Wrangle severity and NDVI data from burned pastures
    reg_dat <-
      BurnIndices  %>%
        group_by(Year, Season, fire) %>%
        summarize(dNBR = mean(dNBR), 
                  NDVI = mean(ndvi), 
                  .groups = 'drop') %>%
        separate(fire, into=c('block','pasture','patch'), sep='-')
  # Fit models
    m0 <- lme4::glmer(dNBR ~ 1 + (1|block) , 
                  family = Gamma(link = "identity"), 
                  data = reg_dat)
    m1 <- lme4::glmer(dNBR ~ NDVI + (1|block), 
                  family = Gamma(link = "identity"), 
                  data = reg_dat)
    m2 <- lme4::glmer(dNBR ~ NDVI + Season + (1|block) , 
                      family = Gamma(link = "identity"), 
                      data = reg_dat)
  # Compare models
    anova(m0, m1, m2)

**Figure 2**

Calculate $z$ scores.

\begin{lstlisting}[language=R]
# NDVI x burn season 
  ndvi_z <-
    LandsatSums %>%
    filter(year %in% c(2017:2020), 
           index == 'ndvi') %>%
    group_by(year, season) %>%
    summarize(Mean = mean(Value), 
              .groups = 'drop') %>%
    left_join(by = 'season', 
              LandsatSums %>%
                filter(index == 'ndvi') %>%
                group_by(season) %>%
                summarize(HistMean = mean(Value, na.rm = T),
                          HistSD = sd(Value, na.rm = T),
                          .groups = 'drop') ) %>%
    mutate(z = (Mean - HistMean)/HistSD, 
           interp = case_when(
             z < -1 ~ "Below", 
             z > 1 ~ "Above", 
             TRUE ~ 'Within' ) )
# YTD precip x burn season 
  prcp <- 
    WxData$Rainfall %>% 
    filter(month(date) %in% c(04:09)) %>%
    separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
    group_by(Year) %>%
    mutate(CumPrecip = cumsum(Rainfall), 
           Month = recode(Month, '05' ="May", '08'='Aug'), 
           Month = paste0(Month, ' ', Day)) %>%
    ungroup() %>%
    filter(Month %in% c('May 15', 'Aug 01' )) %>%
    mutate(season = recode(Month, 'May 15' = 'Spring', 'Aug 01' = 'Summer' )) %>%
    select(Year, season, CumPrecip)
  prcp_z <-
    prcp %>%
    filter(as.numeric(Year) %in% c(2017:2020)) %>%
    group_by(Year, season) %>%
    summarize(Mean = mean(CumPrecip), 
              .groups = 'drop') %>%
    left_join(by = 'season', 
              prcp %>%
                group_by(season) %>%
                summarize(HistMean = mean(CumPrecip, na.rm = T),
                          HistSD = sd(CumPrecip, na.rm = T),
                          .groups = 'drop') ) %>%
    mutate(z = (Mean - HistMean)/HistSD, 
           interp = case_when(
             z < -1 ~ "Below", 
             z > 1 ~ "Above", 
             TRUE ~ 'Within' ))

**Figure 3**

Compare 1990-1999 period to 2012-2022 for fuel and fire weather.

# test historical fuel and fire weather across time periods 
    comp_tab <-
      hist_wx %>%
        group_by(season, variable) %>%
        group_modify(~ broom::tidy(lm(Value ~ period, data = .x))) %>%
        ungroup() %>%
        filter(term != '(Intercept)') %>%
        mutate(P = case_when(
          p.value > 0.05 ~ 'P > 0.05', 
          between(p.value, 0.00099, 0.01) ~ 'P < 0.01', 
          p.value < 0.001 ~ 'P < 0.001', 
          TRUE ~ paste0('P = ', round(p.value, 2)) ), 
          result = paste0('t = ', signif(statistic, 2), ", ", P)) %>%
        select(variable, season, result) %>%
        pivot_wider(names_from = season, 
                    values_from = result) 
\end{lstlisting}

Calculate $z$ scores.

 hist_wx <-
    WxData$HourlyWx %>%
      filter(! date %in% RainyDays$date, 
             ! is.na(season)) %>%
      group_by(Year, date, season) %>%
      summarise_at(.vars = vars(c(DewPoint, RelHum, WindSpeed,VPD)),
                   .funs = c("mean")) %>%
      ungroup() %>%
      pivot_longer(names_to = 'variable', 
                   values_to = 'value', 
                   cols = DewPoint:VPD) %>%
      group_by(Year, season, variable) %>%
      summarise(Value = mean(value), 
                .groups = 'drop') %>%
      mutate(Year = as.character(Year), 
             #season = fct_rev(season), 
             period = case_when(
               Year %in% c(1990:1999) ~  '1990-1999', 
               Year %in% c(2012:2022) ~  '2012-2022', 
               TRUE ~ NA)) %>%
      filter(!is.na(period)) %>%
      bind_rows(WxData$Rainfall %>% 
                  filter(month(date) %in% c(04:09)) %>%
                  separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
                  group_by(Year) %>%
                  mutate(CumPrecip = cumsum(Rainfall), 
                         Month = recode(Month, '05' ="May", '08'='Aug'), 
                         season = paste0(Month, ' ', Day)) %>%
                  ungroup() %>%
                  filter(season %in% c('May 15', 'Aug 01' )) %>%
                  mutate(season = ifelse(season == 'May 15', 'Spring', 'Summer'), 
                         period = case_when(
                           Year %in% c(1990:1999) ~  '1990-1999', 
                           Year %in% c(2012:2022) ~  '2012-2022', 
                           TRUE ~ NA ) )  %>%
                  filter(!is.na(period)) %>%
                  select(Year, season, CumPrecip, period) %>%
                  pivot_longer(names_to = 'variable', 
                               values_to = 'Value', 
                               cols = CumPrecip))  %>%
      bind_rows(HistoricalSpectralData  %>% 
                  select(-pasture) %>%
                  filter(index == 'NDVI', 
                         range != '1990-2022') %>%
                  rename(Year = year, period = range, variable = index) ) 
    hist_wx %>%
      filter(as.numeric(Year) %in% c(2017:2020)) %>%
      group_by(Year, season, variable) %>%
      slice(1) %>%
      summarize(Mean = mean(Value), 
                .groups = 'drop') %>%
      left_join(by = c('season', 'variable'), 
                relationship = "many-to-many",
                hist_wx %>%
                  group_by(period, season, variable) %>%
                  summarize(HistMean = mean(Value, na.rm = T),
                            HistSD = sd(Value, na.rm = T),
                            .groups = 'drop') ) %>%
      mutate(z = (Mean - HistMean)/HistSD, 
             interp = case_when(
               z < -1 ~ "Below", 
               z > 1 ~ "Above", 
               TRUE ~ 'Within') )

### 2025-04-21 SpringSummerSeverity/paper/SupplementalInformation.tex
Spring/summer seasonality R1 update

---
title: Evaluating an attempt to restore summer fire in the Northern Great Plains
author:
- Devan Allen McGranahan and Jay P. Angerer
subtitle:  'Supplementary Information'
header-includes:
  - \usepackage{amsmath}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage[colorlinks=true, citecolor=blue, urlcolor=blue,linkcolor=blue, pdfborder={0 0 0}]{hyperref}
  - \urlstyle{same}
  - \usepackage{xcolor} 
  - \usepackage{booktabs}
  - \raggedright
  - \usepackage{listings}
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
    toc: false
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE)
pacman::p_load(tidyverse, xtable, wesanderson)
  BuGnPal = c( "#67A9CF" ,"#3690C0" ,"#02818A", "#016450")
  FFgrad = wes_palette("FantasticFox1")[c(3,2,4,5)]
  load('../data/NoFireIndices.Rdata')
  # load('../data/BurnIndices.Rdata')
  # load('../data/NoFireTrends.Rdata')
  load('../data/WxData.Rdata')
# some necessary things
  sys_yr <- 
    Sys.Date() %>%
      as.POSIXct(., '%Y-%M-%d') %>%
      year() 
  seasons <- 
    tibble(season = c('Spring', 'Summer'), 
           start = c(as.Date(paste0(sys_yr, '-04-25')), as.Date(paste0(sys_yr, '-08-07'))), 
           end = c(as.Date(paste0(sys_yr, '-05-25')), as.Date(paste0(sys_yr, '-08-27'))))
```

# Additional data and results

## Seasonal trends

```{r ndvi_trend, eval = TRUE, fig.cap="Seasonal trends in Landsat-derived NDVI for the study period. Green shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 4, fig.width = 8}
NoFireIndices %>%
  group_by(ImageDate, pasture) %>%
    summarise(Mean = mean(ndvi), 
              SEM = sd(ndvi)/sqrt(n()), 
              .groups = 'drop') %>%
  mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
         Year = lubridate::year(ImageDate), 
         Year = as.factor(Year), 
         SampleDate = format(ImageDate, '%m-%d'), 
         SampleDate = as.Date(SampleDate, '%m-%d') ) %>%
  filter(Year != 2016) %>%
ggplot(aes(x = SampleDate, 
           color = Year, 
           fill = Year, 
           group = interaction(pasture, Year))) + theme_bw(14) +
  geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf), 
              fill='#A1D99B', alpha=0.5) +
  geom_smooth(aes(y = Mean, group = Year),
              se = T,
              level = 0.99) +
  geom_errorbar(aes(ymin = Mean - SEM, 
                    ymax = Mean + SEM),
                alpha = 0.5) +
  geom_line(aes(y = Mean),
             alpha = 0.5) +
  labs(x = 'Image date', y = 'NDVI') + 
    scale_fill_manual(values = FFgrad) + 
  scale_color_manual(values = FFgrad) + 
  scale_x_date( date_breaks = 'month', 
                date_labels = '%b-%d', 
                minor_breaks = NULL) 
```

```{r precip_trend, eval = TRUE, fig.cap="Historical precipitation data for the entire 42-year period. Thick black line shows smoothed 42-yr trend, gray lines individual years, and colored lines study years. Green shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 4, fig.width = 8}
  WxData$Rainfall %>% 
  mutate(day = day + years(1)) %>%
    filter(month(date) %in% c(04:09)) %>%
    separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
    group_by(Year) %>%
    mutate(CumPrecip = cumsum(Rainfall)) %>%
    ggplot(aes(x = day)) + theme_bw(14) +
    geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf),
              fill='#A1D99B', alpha=0.5) +
    geom_line(aes(y = CumPrecip, group = Year),
              color = 'grey40', 
              alpha = 0.5) +
    geom_smooth(aes(y = CumPrecip),
                se = F,
                color = 'black',
                size = 1,
                level = 0.99) +
    geom_line(data = . %>%
                filter(as.numeric(Year) %in% c(2017:2020)) , 
              aes(y = CumPrecip, 
                  color = Year,
                  group = Year),
              size = 2) +
    labs(x = 'Day of year', y = 'Cumulative rainfall (mm)') + 
    scale_fill_manual(values = FFgrad) + 
    scale_color_manual(values = FFgrad) + 
    scale_x_date( date_breaks = 'month', 
                  date_labels = '%b-%d', 
                  minor_breaks = NULL)
```

@@

## $z$ scores for historical data by burn season

```{r historical_z, eval = TRUE, results='asis'}
load('../data/z_tab.Rdata')
 z_tab$hist %>%
  xtable(caption = 'Seasonal deviations in fuel conditions and fire weather for spring and summer burn seasons, relative to two decades (1990-1999, 2012-2022) as presented in Fig. 4. Anomalies determined by $z$ scores (parentheses) and categorized as above, below, or within 1 standard deviation of the mean for the given historical range. 2018 and 2019 were the years in which no summer burns were conducted; all spring burns were conducted in all years') %>%
   print(include.rownames = F, comment = F)
```

@@

# Script

@@

## Remote sensing

### Sentinel-2

This \texttt{EvalScript} can be used in the [Copernicus browser](https://browser.dataspace.copernicus.eu/) as a Custom script to create a Custom visualization that can be exported as a 16-bit \texttt{.tiff} for raster analysis. 
The user must select the area of interest and imagery dates and is responsible for assessing cloudiness.

@@

\clearpage
### Landsat

This script for [Google Earth Engine](https://code.earthengine.google.com/) exports a composited NDVI \texttt{.tiff} file to the user's Google Drive for the area of interest defined as \texttt{geometry} for an eight week period beginning with each date given in \texttt{listDate}. 
The script combines imagery from Landsat missions 5, 7, \& 8 and handles cloud masking.

@@

## Analysis

@@

### Remotely sensed imagery

@@

### Weather data

This script wrangles the weather data downloaded from the [North Dakota Ag Weather Network's Streeter station](https://ndawn.ndsu.nodak.edu/station-info.html?station=48).

@@

### Statistical analysis

@@

**Figure 1**

@@

**Figure 2**

Calculate $z$ scores.

@@

**Figure 3**

@@

Calculate $z$ scores.

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{amsmath}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[colorlinks=true, citecolor=blue, urlcolor=blue,linkcolor=blue, pdfborder={0 0 0}]{hyperref}
\urlstyle{same}
\usepackage{xcolor}
\usepackage{booktabs}
\raggedright
\usepackage{listings}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Evaluating an attempt to restore summer fire in the Northern Great Plains},
  pdfauthor={Devan Allen McGranahan and Jay P. Angerer},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Evaluating an attempt to restore summer fire in the Northern
Great Plains}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supplementary Information}
\author{Devan Allen McGranahan and Jay P. Angerer}
\date{}

\begin{document}
\maketitle

\section{Additional data and results}\label{additional-data-and-results}

\subsection{Seasonal trends}\label{seasonal-trends}

\begin{figure}
\centering
\includegraphics{SupplementalInformation_files/figure-latex/ndvi_trend-1.pdf}
\caption{Seasonal trends in Landsat-derived NDVI for the study period.
Green shaded areas span the first and last dates of completed burns in
the spring and summer.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SupplementalInformation_files/figure-latex/precip_trend-1.pdf}
\caption{Historical precipitation data for the entire 42-year period.
Thick black line shows smoothed 42-yr trend, gray lines individual
years, and colored lines study years. Green shaded areas span the first
and last dates of completed burns in the spring and summer.}
\end{figure}

@@

\subsection{\texorpdfstring{\(z\) scores for historical data by burn
season}{z scores for historical data by burn season}}\label{z-scores-for-historical-data-by-burn-season}

\begin{table}[ht]
\centering
\begin{tabular}{llllll}
  \hline
Variable & Year & Spring 1990-99 & Spring 2012-22 & Summer 1990-99 & Summer 2012-22 \\ 
  \hline
Fuelbed Greenness (NDVI) & 2017 & Above (1.3) & Within (0.96) & Within (0.34) & Within (-0.05) \\ 
    & 2018 & Within (0.15) & Within (-0.87) & Within (0.92) & Within (0.64) \\ 
    & 2019 & Within (-0.54) & Below (-1.96) & Above (1.67) & Above (1.51) \\ 
    & 2020 & Above (1) & Within (0.48) & Within (0.37) & Within (-0.01) \\ 
  Cumulative rainfall & 2017 & Within (-0.63) & Within (-0.89) & Within (-0.89) & Within (-0.91) \\ 
    & 2018 & Within (-0.76) & Below (-1.08) & Above (1.87) & Above (1.38) \\ 
    & 2019 & Within (-0.27) & Within (-0.37) & Above (1.04) & Within (0.69) \\ 
    & 2020 & Within (0.02) & Within (0.06) & Below (-1.16) & Below (-1.13) \\ 
  Dew Point & 2017 & Below (-1.21) & Within (-0.98) & Below (-1.15) & Below (-1.07) \\ 
    & 2018 & Within (-0.05) & Within (0.63) & Within (-0.27) & Within (-0.07) \\ 
    & 2019 & Below (-1.01) & Within (-0.7) & Within (-0.21) & Within (0) \\ 
    & 2020 & Within (0.41) & Above (1.28) & Within (0.9) & Above (1.27) \\ 
  Relative Humidity & 2017 & Within (-0.85) & Within (-0.46) & Within (-0.56) & Within (-0.37) \\ 
    & 2018 & Within (-0.99) & Within (-0.67) & Within (-0.82) & Within (-0.65) \\ 
    & 2019 & Within (-0.48) & Within (0.12) & Within (0.67) & Above (1.01) \\ 
    & 2020 & Within (0.37) & Above (1.43) & Within (0.32) & Within (0.62) \\ 
  Vapor Pressure Deficit & 2017 & Within (0.48) & Within (0.26) & Within (-0.17) & Within (-0.32) \\ 
    & 2018 & Above (1.26) & Above (1.3) & Within (0.91) & Within (0.82) \\ 
    & 2019 & Within (-0.38) & Within (-0.9) & Below (-1.15) & Below (-1.36) \\ 
    & 2020 & Within (-0.62) & Below (-1.21) & Within (-0.13) & Within (-0.28) \\ 
  Wind speed & 2017 & Within (-0.69) & Within (-0.19) & Below (-1.69) & Within (-0.91) \\ 
    & 2018 & Within (-0.97) & Within (-0.42) & Below (-2.06) & Below (-1.31) \\ 
    & 2019 & Below (-1.54) & Within (-0.89) & Below (-1.3) & Within (-0.5) \\ 
    & 2020 & Below (-1.24) & Within (-0.65) & Within (-0.74) & Within (0.1) \\ 
   \hline
\end{tabular}
\caption{Seasonal deviations in fuel conditions and fire weather for spring and summer burn seasons, relative to two decades (1990-1999, 2012-2022) as presented in Fig. 4. Anomalies determined by $z$ scores (parentheses) and categorized as above, below, or within 1 standard deviation of the mean for the given historical range. 2018 and 2019 were the years in which no summer burns were conducted; all spring burns were conducted in all years} 
\end{table}

@@

\section{Script}\label{script}

@@

\subsection{Remote sensing}\label{remote-sensing}

\subsubsection{Sentinel-2}\label{sentinel-2}

This \texttt{EvalScript} can be used in the
\href{https://browser.dataspace.copernicus.eu/}{Copernicus browser} as a
Custom script to create a Custom visualization that can be exported as a
16-bit \texttt{.tiff} for raster analysis. The user must select the area
of interest and imagery dates and is responsible for assessing
cloudiness.

@@

\subsubsection{Landsat}\label{landsat}

This script for \href{https://code.earthengine.google.com/}{Google Earth
Engine} exports a composited NDVI \texttt{.tiff} file to the user's
Google Drive for the area of interest defined as \texttt{geometry} for
an eight week period beginning with each date given in
\texttt{listDate}. The script combines imagery from Landsat missions 5,
7, \& 8 and handles cloud masking.

@@

\subsection{Analysis}\label{analysis}

@@

\subsubsection{Remotely sensed imagery}\label{remotely-sensed-imagery}

@@

\subsubsection{Weather data}\label{weather-data}

This script wrangles the weather data downloaded from the
\href{https://ndawn.ndsu.nodak.edu/station-info.html?station=48}{North
Dakota Ag Weather Network's Streeter station}.

@@

\subsubsection{Statistical analysis}\label{statistical-analysis}

@@

\textbf{Figure 1}

@@

\textbf{Figure 2}

Calculate \(z\) scores.

@@

\textbf{Figure 3}

@@

Calculate \(z\) scores.

@@

\end{document}
